### 2015-07-06 narrative/data-description.tex

\paragraph{Proposed research approach:} 
We will design and develop new techniques for data management, description,
and access that allow users to describe the data utility based on their
expectations. This can be done by allowing users to plug in functions that
partition user level data into chunks, which are then executed by the middleware to
calculate the utility value for each chunk. This value, along with other
metadata, will be encapsulated by the middleware and utilized by the underlying
storage system to manage data placement.
%
%We will design an updated I/O API that will expand on the current POSIX I/O
%semantics by adding new information to the I/O request.

@@

We will also develop a new type of querying system that allows an application
to ask the storage system about completion timing information. The response
would be an estimated elapsed time for the described operation. We will explore
how these query functions can be integrated into common applications with
minimal code disruptions and the set of policies that applications can use to
respond to this new information. We will also explore how applications can
query the storage system to gain insights on available space in different tiers
and how applications can tailor their output data to match the available space.
Finally we will study the use of an external data annotation system that can
provide information to the storage system without requiring the application to
be recompiled.

%
From an application level, we will explore these techniques through our ADIOS
framework using many of the applications to which we have access, including
XGC1, GTC, and SPECFM3D. We will understand: 1) what hints the users can give
us, 2) how the users can take away much of their application specific code to
``system level code'' through the use of these hints.  We will design an API
that will integrate through the storage layer to give time estimations so that
users can decide which data priorities will be written and read. The utility
value for each chunk will place it into one of ``n'' bins where the first bin
is the most important data and the last bin is the least important data. The
bin will give it a lifetime in the highest level of storage that the user thinks
will be needed. For example: in Figure~\ref{fig:ssio-bucket} we have data that
comes from the XGC1 code.  The user can pick a data-refactoring scheme (see
Sec~\ref{sec:data-refactor}) which will bucket the particle data into 4
prioritization levels.  The user can then pick a data utility, such as 3 months
for the highest priority items and 12 months for the lowest priority data.
The user will then call an API when they specify all of the chunks they want
written (particles, two fields, and a mesh in this example) and the system
will return an estimate of the time it will take to write. The user can then
use another API call which will determine the lowest priority element (in this
case the user will write all of the data (P1-P4). The data will then be stored
on all of the layers of the file system where the two highest priority buckets
will be stored in the Parallel File System, but the next highest priority items
will be stored lower in the storage stack. 
%

>>>>

\paragraph{Proposed research approach:} 
We will design and develop new techniques for data management, description,
and access that allow users to describe the data utility based on their
expectations. This can be done by allowing users to plug in functions that
provide data-specific functionality to the middleware and storage system.
These functions may be executed by the system to
calculate the utility value for each chunk, handle custom compression
and decompression, and perform other data-specific services, thus allowing
the system to more effectively manage data placement and retrieval.
%
%We will design an updated I/O API that will expand on the current POSIX I/O
%semantics by adding new information to the I/O request.

@@

We will also develop a new type of querying system that allows an application
to ask the storage system about completion timing information. The response
would be an estimated time for the described operation. We will explore
how these query functions can be integrated into common applications with
minimal code disruptions and the set of policies that applications can use to
respond to this new information. We will also explore how applications can
query the storage system to gain insights on available space in different tiers
and how applications can tailor their output data to most effectively use the available space.
Finally we will study the use of an external data annotation system that can
provide information to the storage system without requiring the application to
be recompiled.

%
From an application level, we will explore these techniques through our ADIOS
framework using many of the applications to which we have access, including
XGC1, GTC, and SPECFM3D. We will investigate: 1) what hints the users can give
the system, 2) how the users can replace much of their
application specific code with
``system level code'' through the use of these hints.  We will design an API
that will be integrated with the storage layer to give time estimations so that
users can decide which data priorities will be written and read. The utility
value for each chunk will be used to sort the chunks into prioritized bins.
Chunks in a particular bin will be assigned an appropriate storage
 tier and lifetime, which can be decided by the user through the API.
For example: in Figure~\ref{fig:ssio-bucket} we have data that
comes from the XGC1 code.  The user can pick a data-refactoring scheme (see
Sec~\ref{sec:data-refactor}) which will bucket the particle data into 4
prioritization levels.  The user can then pick a data utility, such as 3 months
for the highest priority items and 12 months for the lowest priority data.
The user will then call a function, provided by the API, that allows them
to specify all of the chunks they want
written (particles, two fields, and a mesh in this example) and the system
will return an estimate of the time it will take to write. The user can then
use another call which will determine the lowest priority element (in this
case the user will write all of the data (P1-P4). The data will then be stored
across the layers of the file system where the two highest priority buckets
will be stored in the Parallel File System, but the next highest priority items
will be stored lower in the storage stack. 
%
